# Tidyverse tools

```{r}
library(tidyverse)
```

## Select

### Original Data Set

```{r}
head(storms)



```

You can use (data source, columns)

```{r}
dplyr::select(storms, name, pressure)
```

We can also do (data source, -columns) to choose all but that column, i.e.:

```{r}
dplyr::select(storms, -name)
```

### Other select functions:

+-------------------+----------------------------------------------------------+
| Function call     | Description                                              |
+===================+==========================================================+
| \-                | select everything but                                    |
+-------------------+----------------------------------------------------------+
| :                 | select range                                             |
+-------------------+----------------------------------------------------------+
| contains()        | Select columns whose name contains a character string    |
+-------------------+----------------------------------------------------------+
| ends_with()       | Select columns whose name ends with a string             |
+-------------------+----------------------------------------------------------+
| everything()      | Select every column                                      |
+-------------------+----------------------------------------------------------+
| matches()         | Select columns who name matches a regular expression     |
+-------------------+----------------------------------------------------------+
| num_range()       | Select columns named x1,x2,x3...                         |
+-------------------+----------------------------------------------------------+
| one_of()          | Select columns whose names are in a group of names       |
+-------------------+----------------------------------------------------------+
| starts_with()     | Select columns whose name starts with a character string |
+-------------------+----------------------------------------------------------+

: Other useful selection function

## Filter

This will apply a test to every row in the data frame and return just the rows that pass the test. You can combine by putting a comma in the state. It acts as AND.

```{r}
dplyr::filter(storms, wind >= 50,
              name %in% c("alberto", "Alex", "allison"))
```

### Logical Tests in R

The columns on the right show Boolean operator, these combine 2 or more logical tests into a single one, so you get true or false. The columns on the left actually do logical operations.

+--------------+-----------------------+------------------+----------------+
| Logical Test | ?Comparison           | Boolean Operator | ?base::Logic   |
+==============+=======================+==================+================+
| \<           |                       | &                | boolean and    |
+--------------+-----------------------+------------------+----------------+
| \>           | Greater than          | \|               | boolean or     |
+--------------+-----------------------+------------------+----------------+
| ==           | Equal to              | xor              | exactly or     |
+--------------+-----------------------+------------------+----------------+
| \<=          | Less than or equal    | !                | not            |
+--------------+-----------------------+------------------+----------------+
| \>=          | Greater than or equal | any              | any true       |
+--------------+-----------------------+------------------+----------------+
| %in%         | Group membership      | all              | all true       |
+--------------+-----------------------+------------------+----------------+
| !=           | Not equal             |                  |                |
+--------------+-----------------------+------------------+----------------+
| is.na        | is NA                 |                  |                |
+--------------+-----------------------+------------------+----------------+
| !is.na       | Is not NA             |                  |                |
+--------------+-----------------------+------------------+----------------+

: Logical tests in R

## Mutate

The mutate function takes your data frame and it returns a copy of the data with, with additional variables that you derive from the variable set that exists in the data. So mutate is there for anything you can derive from the data you already have, and want to make a new column from it.

```{r}
storms %>% 
  mutate(ratio = pressure / wind) %>% 
  select(name, pressure, wind, ratio)
```

You can also make multiple columns at the same time, even using columns that are created at the same time, as long as those come before chronologically, i.e.

```{r}
storms %>% 
  mutate(ratio = pressure / wind, inverse = ratio^-1) %>% 
  select(name, pressure, wind, ratio, inverse)
```

One thing to note: mutate doesn't affect the original data frame. It simply returns a new temporary dataframe. If you want to save the columns you need to assign it to a new data frame (or overwrite the original)

### Useful mutate functions:

+--------------------------------------------------------+--------------------------------------------+
| Function                                               | Description                                |
+========================================================+============================================+
| pmin(), pmax()                                         | Element wise min and max                   |
+--------------------------------------------------------+--------------------------------------------+
| cummin(), cummax()                                     | cumulative min and max                     |
+--------------------------------------------------------+--------------------------------------------+
| cumsum(), cumprod()                                    | Cumulative sum and product                 |
+--------------------------------------------------------+--------------------------------------------+
| between()                                              | Are values between a and b?                |
+--------------------------------------------------------+--------------------------------------------+
| cume_dist()                                            | Cumulative distribution of values          |
+--------------------------------------------------------+--------------------------------------------+
| cumall(), cumany()                                     | Cumulative all and any                     |
+--------------------------------------------------------+--------------------------------------------+
| cummean()                                              | Cumulative mean                            |
+--------------------------------------------------------+--------------------------------------------+
| lead(), lag()                                          | Copy with values one position              |
+--------------------------------------------------------+--------------------------------------------+
| ntile()                                                | Bin vector into n buckets                  |
+--------------------------------------------------------+--------------------------------------------+
| dense_rank(), min_rank(), percent_rank(), row_number() | Various ranking methods                    |
+--------------------------------------------------------+--------------------------------------------+

## Summarise / Summarize

Summarize allows us to take a data frame and calculate a summary statistic from it and get back a new data frame that is much smaller.

```{r}
storms %>% 
  summarize(median = median(pressure), variance = var(pressure), n = n())
```

Works very similar to mutate.

### Useful summary functions

+------------------+---------------------------------------------------+
| Function         | Description                                       |
+==================+===================================================+
| min(), max()     | Minimum and maximum values                        |
+------------------+---------------------------------------------------+
| mean()           | Mean value                                        |
+------------------+---------------------------------------------------+
| median()         | Median Value                                      |
+------------------+---------------------------------------------------+
| sum()            | Sum of values                                     |
+------------------+---------------------------------------------------+
| var, sd()        | Variance and standard deviation of a vector       |
+------------------+---------------------------------------------------+
| first()          | First value in a vector                           |
+------------------+---------------------------------------------------+
| last()           | Last value in a vector                            |
+------------------+---------------------------------------------------+
| nth              | Nth value in a vector                             |
+------------------+---------------------------------------------------+
| n()              | The number of values in a vector                  |
+------------------+---------------------------------------------------+
| n_distinct()     | The number of distinct values in a vector         |
+------------------+---------------------------------------------------+

: Useful summary functions, top 5 are specific to dplyr
